IN FOCUS \\ Natural history matters: how biological constraints shape diversified interactions in pollination networks

### 2016-07-10 ms_infocus.tex
Typeset final version for submission.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS: In Focus comment paper for Journal of Animal Ecology.
%% Draft
%% April 2016
%% Revised version: 
%% Original version wordcount: 1716 words.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper,12pt]{article}
\usepackage{comment}
\usepackage{jae}
\title{IN FOCUS \\ Natural history matters: how biological constraints shape diversified interactions in pollination networks}
\running{Forbidden interactions}

@@

\nwords{1716}
\ntables{0}
\nfig{2}
\nref{14}

@@

%---------------------------------------------------------------------------------
Size-limited foragers show clear restrictions on the size of prey items they can efficiently handle. In the case of plant-pollinator interactions, size uncoupling between pollinator bodies and flower sizes or structure are specially relevant in filtering out a range of potential partners \citep{Cocucci:2009}. The idea, when applied to the bizarre flowers of some plants pollinated by sphingid moths (Lepidoptera: Sphingidae) (Fig. 1), was seminal in Darwinian evolutionary theory to support the potential of natural selection in shaping adaptations \citep{Arditti:2012}. Wallace \citep{Wallace:1867} in his book, \textit{Creation by law}, vividly uses the famous example of the Malagasy orchid and its sphingid pollinator to refute the arguments of the Duke of Argyll against natural selection and Darwinism:

@@

 Phenotypic fitting of corolla length and shape and the pollinators' feeding apparatus and body sizes are important because the better the fit, the better the consequences in terms of fitness outcomes for the interaction partners \citep{Nilsson:1988}. Yet the expectation of perfect trait matching across populations or communities is too simplistic \citep{Anderson:2010}: "arms races" as initially suggested by Darwin and Wallace are frequently asymmetric, originating pollinator shifts rather than tight phenotypic trait matching (Fig. 2)\citep{Wasserthal:2014}. Extensive local variation in phenotypic mismatch exists in different plant-pollinator systems \citep[e.g., ][]{Cocucci:2009,Anderson:2010,More:2012}, with pollinator-mediated selection geographic mosaics of locally coevolved partners.

Recent work by Sazatornil \textit{et al.} \citep{Sazatornil:2016} provides evidences that the types of trait mismatching outlined in Fig. 2 limit the ranges of host plants for sphingid pollinators, and ultimately shape their complex plant-pollinator networks. By using a comparative analysis of five different hawkmoth/flower assemblages across four South American biotas (Atlantic rainforest and Cerrado in Brazil, Chaco montane dry woodland, and the ecotone between western Chaco woodland and Yungas montane rain forest in Argentina) they tested the contributions of phenotypic matching to explain observed patterns of moth-flower interactions.

How are these moth-flower interactions assembled? Sazatornil \textit{et al.} \citep{Sazatornil:2016} first tested a neutral model, where interactions are independent of trait-matching. Under this hypothesis distribution parameters (mean and standard deviation) must be the same for both distributions. They further tested a Forbidden links hypothesis, where interactions occurred only if the hawkmoth proboscis length (HPL) is equal to or greater than the effective length of the flower (EFL). EFL is just the corolla tube length (as in Fig. 2 for long-tubed and salverform corollas) or the stamen protrusion length in brush-type and funnel-shape flowers (as in Fig. 1). Sazatornil \textit{et al.} further tested the morphological match hypothesis, where the probability of occurrence of an interaction depends on the frequency of possible pairwise differences between HPL and EFL, i.e., all possible pairwise HPL-EFL differences were weighted by their respective interaction frequency.

Th trait matching between HPL and EFL is crucial in this type of interaction and determines its outcome in terms of fitness for both partners. Nilsson \citep{Nilsson:1988} demonstrated experimentally that shortening the nectary tube of long-spurred corollas decreased both seed set and pollinia removal for \textit{Platanthera} orchids. Further experimental evidence has been provided for long-tongued nemestrinid flies pollinating long-tubed irises in South Africa, where increased mismatch decreases both plant fitness and the nectar extraction efficiency of the pollinators \citep{Pauw:2009}. Sazatornil \textit{et al.} extend those results to the scale of the whole moth-plant assemblage and demonstrate that trait matching successfully predicts the diversity of interactions recorded. Interestingly enough, the interaction patterns in two local assemblages from ecotone areas of the Chaco woodland-Yungas montane rain forest transition are better fitted by a neutral model where pairwise interactions are driven by probability of interspecific encounter. Yet Sazatornil \textit{et al.} did not include the morphological difference for parameter estimation when interactions were not recorded. Thus the test of the mismatch hypothesis implicitly includes forbidden links effects: a full mismatch of corolla tube/proboscis lengths actually means a forbidden link. Furthermore,  a fraction of unobserved interactions was likely caused by phenological uncoupling between flowering and hawkmoth activity phenophases \citep{BasJor:2014,Sazatornil:2016}. In any case the mismatch hypothesis somehow captures the fact that a fraction of the unobserved interactions in these hawkmoth/flower assemblages is due to extreme phenotypic mismatch, i.e., size-related forbidden links \citep{Sazatornil:2016}; also see \citep{VizentinBugoni:2014} for evidences with hummingbird-flower interactions.

% Forbidden links background.
% Explain the biological basis of mismatch using Fig 2 and Arditti
Forbidden links represent a family of causes for not observing specific interactions when sampling diversified plant-animal interaction networks, and stem on biological causes deeply linked to the fascinating natural history details of these interactions \citep{BasJor:2014}. The raw material for phenotypic mismatches in the specific case of hawkmoth-flower interactions is the extreme variability of the two pivotal traits determining their outcomes: proboscis length and corolla/spur or nectary depth (Fig. 2) \citep{Cocucci:2009,Arditti:2012,Nilsson:1988,Miller:1997}.

Sazatornil et al. approach would be most useful for proper tests of coevolutionary hypotheses in hawkmoth/flower assemblages (and plant-animal mutualisms in general): assessing match/mismatch patterns for every possible pairwise interaction among partners within complex webs of interaction where multiple life-history attributes may contribute biological reasons to expect forbidden links. The morphological match hypothesis is not the only mechanism to explain patterns of hawkmoth–plant interactions, where other life-history limitations may operate generating forbidden links, e.g., phenological mismatches (for example in the case of long-distance or elevational migratory hawkmoths), constraints from foraging for oviposition sites \citep{Alarcon:2008}, energetic constraints due to balances of nectar availability/foraging costs \citep{Borrell:2005}, etc.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS: In Focus comment paper for Journal of Animal Ecology.
%% Draft
%% April 2016
%% Revised version: 
%% Original version wordcount: 1928 words.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper,12pt]{article}
\usepackage{comment}
\usepackage{jae}
\title{IN FOCUS \\ Natural history matters: how biological constraints shape diversified interactions in pollination networks}
\running{Forbidden interactions}

@@

\nwords{1928}
\ntables{0}
\nfig{2}
\nref{14}

@@

%---------------------------------------------------------------------------------
Size-limited foragers show clear restrictions on the size of prey items they can efficiently handle. In the case of plant-pollinator interactions, size uncoupling between pollinator bodies and flower sizes (i.e., length of spur or corolla tubes) or structure are specially relevant in filtering out a range of potential partners \citep{Cocucci:2009}. As a general trend, also observed in frugivorus animals dispersing seeds, larger (long-tongued) species can exploit longer-tubed flowers and a wider range of flower sizes. The size of the feeding apparatus (e.g., tongue length, gape width) limits the maximum size of flower or fruit an animal mutualist can efficiently use. Phenotypic trait matching is thus a key influence in the effectiveness of plant-animal interactions shaped by these size effects, where the interaction outcomes depend on close matching. The idea, when applied to the bizarre flowers of some plants pollinated by sphingid moths (Lepidoptera: Sphingidae) (Fig. 1), was seminal in Darwinian evolutionary theory to support the potential of natural selection in shaping adaptations \citep{Arditti:2012}. Wallace \citep{Wallace:1867} in his book, \textit{Creation by law}, vividly uses the famous example of the Malagasy orchid and its sphingid pollinator to refute the arguments of the Duke of Argyll against natural selection and Darwinism:

@@

 Phenotypic fitting of corolla length and shape and the pollinators' feeding apparatus and body sizes are important because the better the fit, the better the consequences in terms of fitness outcomes for the interaction partners \citep{Nilsson:1988}. Yet the expectation of perfect trait matching across populations or communities is too simplistic \citep{Anderson:2010}: "arms races" as initially suggested by Darwin and Wallace are frequently asymmetric, originating pollinator shifts rather than tight phenotypic trait matching (Fig. 2)\citep{Wasserthal:2014}. Therefore, extensive local variation in phenotypic mismatch exists in different plant-pollinator systems \citep[e.g., ][]{Cocucci:2009,Anderson:2010,More:2012}, with pollinator-mediated selection geographic mosaics of locally coevolved partners where tight phenotypic matching is not necessarily the rule.

Recent work by Sazatornil \textit{et al.} \citep{Sazatornil:2016} provides evidences that the types of trait mismatching outlined in Fig. 2 limit the ranges of host plants for sphingid pollinators, and ultimately shape their complex plant-pollinator networks. By using a comparative analysis of five different hawkmoth/flower assemblages across four South American biotas (Atlantic rainforest and Cerrado in Brazil, Chaco montane dry woodland, and the ecotone between western Chaco woodland and Yungas montane rain forest in Argentina) they tested the contributions of phenotypic matching to explain observed patterns of moth-flower interactions. Pairwise patterns of trait matching scale-up to conform complex webs of interaction where in some cases interaction topology is linked to phenotypic fit between partners while in others the interaction pattern appears more determined by neutral processes (i.e., probability of interspecific encounter).

How are these moth-flower interactions assembled? Sazatornil \textit{et al.} \citep{Sazatornil:2016} use a comparative analysis of hawkmoth proboscis length ($HPL$) and the effective length of the flower ($EFL$) for every pairwise interaction recorded. EFL is just the corolla tube length (as in Fig. 2 for long-tubed and salverform corollas) or the stamen protrusion length in brush-type and funnel-shape flowers (as in Fig. 1). The authors first tested a neutral model, where interactions are independent of trait-matching and driven solely by variation in local abundance. Under this hypothesis distribution parameters (mean and standard deviation) must be the same for both the simulated and empirical distributions. They further tested a Forbidden links hypothesis, where interactions occurred only if $HPL \geq EFL$. Sazatornil \textit{et al.} further tested the morphological match hypothesis, where the probability of occurrence of an interaction depends on the frequency of possible pairwise differences between $HPL$ and $EFL$, i.e., all possible pairwise $HPL-EFL$ differences were weighted by their respective interaction frequency.

Th trait matching between $HPL$ and $EFL$ is crucial in this type of interaction and determines its outcome in terms of fitness for both partners. Nilsson \citep{Nilsson:1988} demonstrated experimentally that shortening the nectary tube of long-spurred corollas decreased both seed set and pollinia removal for \textit{Platanthera} orchids. Further experimental evidence has been provided for long-tongued nemestrinid flies pollinating long-tubed irises in South Africa, where increased mismatch decreases both plant fitness and the nectar extraction efficiency of the pollinators \citep{Pauw:2009}. Sazatornil \textit{et al.} extend those results to the scale of the whole moth-plant assemblage and demonstrate that trait matching successfully predicts the diversity of interactions recorded. Interestingly enough, the interaction patterns in two local assemblages from ecotone areas of the argentinian Chaco woodland-Yungas montane rain forest transition are better fitted by a neutral model where pairwise interactions are driven by probability of interspecific encounter. Yet Sazatornil \textit{et al.} did not include the morphological difference for parameter estimation when interactions were not recorded. Thus the test of the mismatch hypothesis implicitly includes forbidden links effects: a full mismatch of corolla tube/proboscis lengths actually means a forbidden link. Furthermore, a fraction of unobserved interactions was likely caused by phenological uncoupling between flowering and hawkmoth activity phenophases \citep{BasJor:2014,Sazatornil:2016}. In any case the mismatch hypothesis somehow captures the fact that a fraction of the unobserved interactions in these hawkmoth/flower assemblages is due to extreme phenotypic mismatch, i.e., size-related forbidden links \citep{Sazatornil:2016}; also see \citep{VizentinBugoni:2014} for evidences with hummingbird-flower interactions.

% Forbidden links background.
% Explain the biological basis of mismatch using Fig 2 and Arditti
Forbidden links represent a family of causes for not observing specific interactions when sampling diversified plant-animal interaction networks, and stem on biological causes deeply linked to the fascinating natural history details of these interactions \citep{BasJor:2014}. The raw material for phenotypic mismatches in the specific case of hawkmoth-flower interactions is the extreme variability of the two pivotal traits determining their outcomes: proboscis length and corolla/spur or nectary depth (Fig. 2) \citep{Cocucci:2009,Arditti:2012,Nilsson:1988,Miller:1997}. This variation is extensive for both the plant and pollinator partners and observable at individual, population and species levels.

Sazatornil et al. approach would be most useful for proper tests of coevolutionary hypotheses in hawkmoth/flower assemblages (and plant-animal mutualisms in general): assessing match/mismatch patterns for every possible pairwise interaction among partners within complex webs of interaction where multiple life-history attributes may contribute biological reasons for forbidden links. The morphological match hypothesis is not the only mechanism to explain patterns of hawkmoth–plant interactions, where other life-history limitations may operate generating forbidden links, e.g., phenological mismatches (for example in the case of long-distance or elevational migratory hawkmoths), constraints from foraging for oviposition sites \citep{Alarcon:2008}, energetic constraints due to balances of nectar availability/foraging costs \citep{Borrell:2005}, etc. Most importantly, this paper nicely illustrates how the fascinating natural history details of these interactions can be use to build meaningful testable models to assess the mechanisms beyond structure and function of megadiversified webs of interactions among free-living species.
